Geometric models overestimate lake depth due to imperfect slope measurement

### 2020-11-05 manuscript/agujournaltemplate.tex
fix internal figure numbering

Although proxy distance to lake center was often much larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8). In contrast, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17). For slope measures, most lakes had lower magnitude (i.e. shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig3}). Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope. In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes. For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4). However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S4).

\begin{figure}
  \begin{center}
    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
  \end{center}
  \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope. A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig3}
\end{figure}

The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table2}). More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m). Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig4}). We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig4}).

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
  \end{center}
  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig4}
\end{figure}

>>>>

Although proxy distance to lake center was often much larger in magnitude compared to the true distance to the deepest point of lakes’, they were strongly related ($R^2$ = 0.8). In contrast, proxy nearshore land slope and true in-lake slope were more weakly related ($R^2$ = 0.17). For slope measures, most lakes had lower magnitude (i.e. shallower) nearshore land slope compared to true in-lake slope (Figure \ref{fig2}). Taken together, these results suggest that proxy distance to the center of lakes is representative of true distance to the deepest point of lakes whereas proxy nearshore land slope is not representative of true in-lake slope. In addition to overall differences between slope and distance measures, we found differences in these relationships among lake shape classes. For example, in-lake slope and distance to the deepest point of the lake metrics were consistently larger in magnitude for convex lakes as compared to concave lakes (Figure S4). However,  there were not similar differences among slope and distance metrics for natural lakes versus reservoirs (Figure S4).

\begin{figure}
  \begin{center}
    \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/01_geometry_base-1}
  \end{center}
  \caption{Comparison among proxy and true values of lake geometry for A) distance to deepest point versus distance distance to lake center and B) nearshore land slope versus in-lake slope. A best-fit line and coefficient of determination is shown to illustrate representativeness.}\label{fig2}
\end{figure}

The use of proxy nearshore land slope had a larger effect on model fit and prediction error than the use of proxy distance to lake center (Table \ref{table2}). More specifically, the true slope - proxy distance model had a better fit ($R^2$ = 0.73) and lower prediction error (RMSE = 4.23m) compared to the proxy slope - true distance model ($R^2$ = 0.26, RMSE = 6.87m). Furthermore, analysis of model residuals showed overestimation of lake depth for concave lakes when models included a proxy slope measure (Figure \ref{fig3}). We observed similar but smaller overestimation depending on if a lake was classified as a reservoir rather than a natural lake (Figure \ref{fig3}).

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{../figures/02_depth_model_grid_resid-1}
  \end{center}
  \caption{Depth model residuals (residual  = observed - predicted) in meters by cross-section shape and reservoir class indicating overprediction of concave and reservoir lakes.}\label{fig3}
\end{figure}
